Clone and graft:\\ Testing scientific applications as they are built

### 2014-12-19 turcksin-spring-mechanics/paper.tex
Add another section.

Let's implement a program that can solve these questions. Because we want it
to be simple and accessible, we're going to use Python as it's widely used and
because it comes with a number of tools that are going to make our life
simpler by keeping the program short.%
\footnote{In order to run this program, you will have to have the Python
  NumPy, SciPy, and MatPlotLib packages installed on your system, along with
  the basic Python interpreter.}
Here it is:
\begin{lstlisting}[frame=single,basicstyle=\footnotesize]
  import json
  ...
\end{lstlisting}
The program reads its input parameters from a file in JSON (see
\url{http://en.wikipedia.org/wiki/JSON}) format.%
\footnote{JSON is not the prettiest format to write input files in and we
  don't recommend you use it for your programs (it is typically intended for
  exchange of data between programs, not between humans and computers). But it
  will serve for our purposes here primarily because Python has a built-in
  parser for it that makes our program pleasantly free of the tedium of
  parsing input files.}
Here is an example that you can find in the file \texttt{testcase-1.json}:
\lstinputlisting[frame=single,basicstyle=\footnotesize]{tests/testcase-1.json}
In JSON, the name of a parameter is to the left of the colon and its value is
to the right; name/value pairs are separated by commas. The value of a
parameter can be a plain number, or it can be an array if enclosed in
brackets. Arrays can also be nested, which is what we use here for the initial
positions and velocities of our two bodies, each of which lives in
three-dimensional space. With this explanation, it is clear that the input file
specifies properties of the two masses as $\mathbf x_{1,0}=(1,2,3)^T, \mathbf
x_{2,0}=(4,5,6)^T, \mathbf v_{1,0}=(0,0,0)^T, \mathbf v_{2,0}=(1,1,1)^T,
m_1=13.5, m_2=29.75$ and $D=42, L=2.25$ for the spring that connects them, 
thereby completely determining everything in equations
\eqref{eq:ode}--\eqref{eq:ic2}.

@@

Implementing all of this leads to the following program:
\begin{lstlisting}[frame=single,basicstyle=\footnotesize]
  ...
\end{lstlisting}

How do we test this? First, we should check whether the old input files still
work:
\begin{lstlisting}[frame=single,basicstyle=\footnotesize,language=csh]
  python masses-and-springs.py testcase-1.json > testcase-1.output.new
  diff testcase-1.output testcase-1.output.new
  python masses-and-springs.py testcase-2.json > testcase-2.output.new
  diff testcase-2.output testcase-2.output.new
\end{lstlisting}
With the code that accompanies this paper, \texttt{diff} reports no difference
between the old and the new output files, so this is reassuring.

@@

\subsection{Dealing with many tests}

@@

>>>>

Let's implement a program that can solve these questions. Because we want it
to be simple and accessible, we're going to use Python as it's widely used and
because it comes with a number of tools that are going to make our life
simpler by keeping the program short.%
\footnote{In order to run this program, you will have to have the Python
  NumPy, SciPy, and MatPlotLib packages installed on your system, along with
  the basic Python interpreter.}
Here it is:
\marginpar{After documenting spring.py, split it and simplify the first
  version to the pieces we need here}
\lstinputlisting[frame=single,basicstyle=\footnotesize,language=Python]{spring.py}
The program reads its input parameters from a file in JSON (see
\url{http://en.wikipedia.org/wiki/JSON}) format.%
\footnote{JSON is not the prettiest format to write input files in and we
  don't recommend you use it for your programs (it is typically intended for
  exchange of data between programs, not between humans and computers). But it
  will serve for our purposes here primarily because Python has a built-in
  parser for it that makes our program pleasantly free of the tedium of
  parsing input files.}
Here is an example that you can find in the file \texttt{tests/testcase-1.json}:
\lstinputlisting[frame=single,basicstyle=\footnotesize]{tests/testcase-1.json}
In JSON, the name of a parameter is to the left of the colon and its value is
to the right; name/value pairs are separated by commas. The value of a
parameter can be a plain number, or it can be an array if enclosed in
brackets. Arrays can also be nested, which is what we use here for the initial
positions and velocities of our two bodies, each of which lives in
three-dimensional space. With this explanation, it is clear that the input file
specifies properties of the two masses as $\mathbf x_{1,0}=(1,2,3)^T, \mathbf
x_{2,0}=(4,5,6)^T, \mathbf v_{1,0}=(0,0,0)^T, \mathbf v_{2,0}=(1,1,1)^T,
m_1=13.5, m_2=29.75$ and $D=42, L=2.25$ for the spring that connects them, 
thereby completely determining everything in equations
\eqref{eq:ode}--\eqref{eq:ic2}.

@@

Implementing all of this leads to the following program:
\lstinputlisting[frame=single,basicstyle=\footnotesize,language=Python]{spring.py}

How do we test this new functionality? First, we should check whether the old input files still
work:
\begin{lstlisting}[frame=single,basicstyle=\footnotesize,language=csh]
  python masses-and-springs.py testcase-1.json > testcase-1.output.new
  diff testcase-1.output testcase-1.output.new
  python masses-and-springs.py testcase-2.json > testcase-2.output.new
  diff testcase-2.output testcase-2.output.new
\end{lstlisting}
With the code that accompanies this paper, \texttt{diff} reports no difference
between the old and the new output files, so this is reassuring.

@@

\subsection{Dealing with many tests}
\label{sec:many-tests}

@@

\subsection{Dealing with overlapping tests}

In an ideal world, every testcase would only check a single feature (such
testcases are then often called \textit{unit tests}). This way, if something
goes wrong with a change to the source code, a single test will fail and it
will be easy to identify what went wrong.

In practice, this is often difficult to achieve. For example, if we change the
implementation of the function that provides $f(t,y)$ above and break it, it
is likely that \textit{all} testcases will start to fail. Similarly, if you
build testcases by cloning and grafting onto existing ones, whenever the old
one fails, the new one will likely do so as well.

From experience, it turns out that this is not often a real problem. The vast
majority of patches concern just small, isolated pieces of the software and in
those cases, only one or a few testcases will start failing. It is then easy
to identify the simplest one of these (assuming that they were given useful
names, see above in Section~\ref{sec:many-tests}), and that is the one one
will then use for debugging. If one breaks fundamental functionality somewhere
in the core of the software, a large number of tests will start to fail
but in those cases you will either be able to quickly find out what is
happening, or be able to debug the problem by choosing \textit{any} simple
test that was among those that broke. In summary, from a practical
perspective, we have found that it is not usually a problem to have testcases
that check overlapping parts of the software.
